\begin{landscape}
\small
\renewcommand{\arraystretch}{0.3}

\begin{longtable}{|p{0.28\linewidth}|p{0.48\linewidth}|p{0.14\linewidth}|}
\caption{Session 4 canonical functions: what to implement and which session 3 block they come from (dataset loaded from CSV).} \\
\hline
\textbf{Function} & \textbf{Code to put inside the function (from Session 3)} & \textbf{Notes} \\
\hline
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline
\textbf{Function} & \textbf{Code to put inside the function (from Session 3)} & \textbf{Notes} \\
\hline
\endhead

\hline \multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
\endfoot

\hline \hline
\endlastfoot

\texttt{make\_print\_status(status\_text)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
print(f"[STATUS] {status_text}")
\end{minted}
& Optional helper for readable progress logs. \\
\hline
\texttt{determine\_binary\_label(sample, settings)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
if sample[settings["feature_name"]] < settings["threshold"]:
	return settings["positive_label"]
return settings["negative_label"]
\end{minted}
& (binary: setosa vs not\_setosa). \\
\hline
\texttt{determine\_true\_binary\_label(sample, settings)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
if sample[settings["label_key"]] == settings["positive_label"]:
	return settings["positive_label"]
return settings["negative_label"]
\end{minted}
& Converts species labels into binary truth labels. \\
\hline
\texttt{evaluate\_prediction(y\_pred, y\_true)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
return (y_pred == y_true)
\end{minted}
& Returns a boolean correctness flag. \\
\hline
\texttt{update\_metrics(metrics, y\_pred, y\_true)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
if y_pred == y_true:
	metrics["correct"] += 1
else:
	metrics["wrong"] += 1
metrics["total"] += 1
metrics["y_pred_list"].append(y_pred)
\end{minted}
& Refactors the counter update + prediction list append block. \\
\hline
\texttt{classify\_sample(sample, settings)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
return determine_binary_label(sample, settings)
\end{minted}
& Keeps the main loop clean and prepares for Session 5/6 changes. \\
\hline
\texttt{load\_iris\_data(filepath)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
dataset = []
with open(filepath, "r", encoding="utf-8") as f:
	reader = csv.DictReader(f)
	for row in reader:
		dataset.append({...})
\end{minted}
& \textbf{Remark:} this session we load the dataset from \texttt{CSV} (not JSON). Convert numeric fields to \texttt{float}. \\
\hline
\texttt{get\_default\_settings()} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
return {
	"threshold": 2.0, "feature_name": "petal_length",
	"positive_label": "setosa", "negative_label": "not_setosa",
	"label_key": "species"
}
\end{minted}
& Keeps Session 2/3 naming stable; no external settings file yet. \\
\hline
\texttt{initialize\_predictions(dataset, settings)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
for sample in dataset:
	y_pred = classify_sample(sample, settings)
	y_true = determine_true_binary_label(sample, settings)
	update_metrics(metrics, y_pred, y_true)
	print(f"id=... | true=... | pred=... | petal_length=...")
\end{minted}
& This is the session 3 loop, but decomposed into functions. \\
\hline
\texttt{setup\_application(filepath)} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
dataset = load_iris_data(filepath)
settings = get_default_settings()
result = initialize_predictions(dataset, settings)
\end{minted}
& Small pipeline wrapper for readability. \\
\hline
\texttt{main()} & \begin{minted}[fontsize=\small, linenos, breaklines, xleftmargin=2em]{python}
print("Correct:", result["correct"])
\end{minted}
& Runs the whole program and prints the final summary. \\
\hline
\end{longtable}
\end{landscape}