\chapter{Positional and Keyword Arguments}

\section{Introduction}

\begin{tcolorbox}[title=Goal]
\textbf{Refactor} the Session 4 classifier so the same prediction steps can be controlled from function calls. In this session, students learn how positional arguments depend on order, how keyword arguments improve readability, and how default values let one function support several run styles.
\end{tcolorbox}

\begin{tcolorbox}[title=You will work in]
Use the following file for this session:
\begin{itemize}
    \item \texttt{session5/session5\_iris\_template.py} (student template)
\end{itemize}
\textbf{How to run the file:} Use the terminal command: \\
\texttt{python3 session5/session5\_iris\_template.py}
\end{tcolorbox}

In Session 4, students moved the classifier into functions such as \texttt{derive\_true\_label(...)}, \texttt{update\_result\_counts(...)}, and \texttt{calculate\_accuracy(...)}.

In Session 5, we keep those same ideas, but we make the function calls more flexible:
\begin{itemize}
    \item we keep the same prediction rule,
    \item we keep the same counters,
    \item and we add better function interfaces with default values and keyword arguments.
\end{itemize}

\begin{tcolorbox}[title=Why this matters]
Session 5 is the bridge between Session 4 and Session 6.
\begin{itemize}
    \item Session 4 organizes the code into separate functions.
    \item Session 5 keeps those same functions, but improves how they are called.
    \item Session 6 turns those same steps into methods on a class.
\end{itemize}
\end{tcolorbox}

\section{What You Will Learn}
\textbf{In this session, you will learn to:}
\begin{enumerate}\setlength{\itemsep}{5pt}
\item \textbf{Use positional arguments carefully}: understand how Python matches values by order.
\item \textbf{Use keyword arguments clearly}: understand how Python matches values by parameter name.
\item \textbf{Combine positional and keyword arguments}: use both styles in one function call.
\item \textbf{Add default values}: make functions easier to call while keeping sensible behavior.
\item \textbf{Use clearer function names}: give a wrapper function a name that matches what it really does.
\item \textbf{Prepare for Session 6}: see how the same steps can later move into a class.
\end{enumerate}

\subsection{Parameters and Arguments}
These two words are often confused by beginners, so we state the difference clearly:
\begin{itemize}
    \item A \textbf{parameter} is the name written in the function definition.
    \item An \textbf{argument} is the real value passed into that parameter when the function is called.
\end{itemize}

\begin{minted}[fontsize=\small, linenos, breaklines=true, frame=lines]{python}
def make_print_status(status_text):
    print(f"[STATUS] {status_text}")

make_print_status("Build dataset")
\end{minted}

In the example above:
\begin{itemize}
    \item \texttt{status\_text} is the parameter,
    \item \texttt{"Build dataset"} is the argument.
\end{itemize}

\subsection{Why Session 5 Matters}
In a real program, the same logic is often reused with slightly different settings. A good function should not force the caller to rewrite the internal code every time.

Session 5 therefore teaches students a second level of refactoring:
\begin{itemize}
    \item Session 4 refactored \textbf{structure} by moving code into functions.
    \item Session 5 refactors \textbf{interfaces} by improving how those functions are called.
\end{itemize}

\section{Session 4 to Session 5 Refactor Map}
The classifier logic stays the same, but the interface becomes more flexible:
\begin{itemize}
    \item Session 4 separate top-level steps $\rightarrow$ Session 5 one wrapper function: \texttt{run\_classifier\_pipeline(threshold=2.0, print\_each=True)}
    \item Session 4 positional accuracy call $\rightarrow$ Session 5 keyword-friendly \texttt{calculate\_accuracy(correct=0, total=0)}
    \item Session 4 positional-only calls $\rightarrow$ Session 5 positional, keyword, and mixed calls
    \item Session 4 helper functions $\rightarrow$ Session 5 keeps them, renames the main wrapper more clearly, and prepares them for Session 6 methods
\end{itemize}

\section{Canonical Function Set}
Complete these functions in \texttt{session5\_iris\_template.py}. The table below shows the Session 5 target structure:

\input{session5/tab5.tex}

\begin{tcolorbox}[title=Good Practice]
This session is intentionally explicit. When students are learning arguments, confusion usually comes from two places:
\begin{itemize}
    \item not knowing which function header changed,
    \item not knowing which call site must also change.
\end{itemize}
That is why the instructions below always name both the task title and the matching comment in the template.
\end{tcolorbox}

\input{session5/ses5_keyword_argument.tex}
\input{session5/ses5_positional_argument.tex}

\section{A Reusable Line-by-Line Debugging Workflow}
\begin{tcolorbox}[title=Line-by-Line Debugging Workflow]
When students are unsure how an argument moves through the program, use this routine:
\begin{enumerate}
    \item Set a breakpoint on the line in \texttt{main()} that calls the function you want to study.
    \item Start the debugger and stop at that line before the function call happens.
    \item Read the argument values from left to right.
    \item Use \textbf{Step Into} to enter the function.
    \item Compare the values at the call site with the parameter names in the function header.
    \item Keep a close eye on variables such as \texttt{threshold}, \texttt{print\_each}, \texttt{y\_pred}, and \texttt{y\_true}.
    \item If the behavior looks wrong, ask: ``Did the value go into the correct parameter?''
\end{enumerate}
\end{tcolorbox}

\section{What Changed from Session 4}

This session keeps the same beginner-friendly classifier logic from Session 4, but changes how that logic is \textbf{called}.

\begin{itemize}
    \item The accuracy formula is still the same, but \texttt{calculate\_accuracy(...)} now uses default values so it can be called with keyword arguments.
    \item The main wrapper function now has a clearer name: \texttt{run\_classifier\_pipeline(...)}. This helps students see that the function does more than setup.
    \item The prediction loop is still the same idea, but \texttt{run\_prediction\_loop(...)} now demonstrates a mixed call: one positional argument plus keyword arguments.
    \item The metric update logic is still the same, but \texttt{update\_result\_counts(...)} is now used as a second example of combining positional and keyword arguments in one call.
    \item The final runs in \texttt{main()} now show four calling styles: default call, positional override, keyword override, and reordered keyword override.
\end{itemize}

\begin{tcolorbox}[title=Why This Matters]
Students should notice that Session 5 is not mainly about changing the classifier rule. It is about learning how to control the same rule more clearly from function calls.
\end{tcolorbox}

\section{Expected Output When Submitting in Itel}

In the terminal, you should see output like the following when running the completed answer file:
\begin{minted}[fontsize=\small, linenos, breaklines=true, frame=lines]{python}
[STATUS] Default run

=== Start Session 5 Prediction Loop ===
id=flower1 | true=setosa | pred=setosa | petal_length=1.4
id=flower2 | true=setosa | pred=setosa | petal_length=1.5
id=flower3 | true=setosa | pred=setosa | petal_length=1.3
id=flower4 | true=not_setosa | pred=not_setosa | petal_length=4.5
id=flower5 | true=not_setosa | pred=not_setosa | petal_length=4.7
id=flower6 | true=not_setosa | pred=not_setosa | petal_length=5.1
id=flower7 | true=not_setosa | pred=not_setosa | petal_length=6.0

=== Session 5 Summary ===
Correct: 7
Wrong: 0
Total: 7
Accuracy (%): 100.0
All predictions: ['setosa', 'setosa', 'setosa', 'not_setosa', 'not_setosa', 'not_setosa', 'not_setosa']
[STATUS] Positional override run

=== Start Session 5 Prediction Loop ===

=== Session 5 Summary ===
Correct: 7
Wrong: 0
Total: 7
Accuracy (%): 100.0
All predictions: ['setosa', 'setosa', 'setosa', 'not_setosa', 'not_setosa', 'not_setosa', 'not_setosa']
[STATUS] Keyword override run

=== Start Session 5 Prediction Loop ===

=== Session 5 Summary ===
Correct: 7
Wrong: 0
Total: 7
Accuracy (%): 100.0
All predictions: ['setosa', 'setosa', 'setosa', 'not_setosa', 'not_setosa', 'not_setosa', 'not_setosa']
[STATUS] Reordered keyword run

=== Start Session 5 Prediction Loop ===

=== Session 5 Summary ===
Correct: 7
Wrong: 0
Total: 7
Accuracy (%): 100.0
All predictions: ['setosa', 'setosa', 'setosa', 'not_setosa', 'not_setosa', 'not_setosa', 'not_setosa']
\end{minted}

\section{Summary and Review}
\begin{itemize}
    \item Positional arguments are matched by order.
    \item Keyword arguments are matched by parameter name.
    \item Positional and keyword arguments can be combined in one call, as long as the positional arguments come first.
    \item Default values let a function stay easy to call.
    \item Session 5 keeps the Session 4 logic, but gives it a clearer calling interface before Session 6 turns it into a class.
\end{itemize}

\textbf{Reflection Questions:}
\begin{enumerate}
    \item Why can positional arguments become harder to read when a function has many parameters?
    \item Why does \texttt{calculate\_accuracy(correct=..., total=...)} make the call site easier to understand?
    \item Why does the mixed call \texttt{run\_prediction\_loop(dataset, threshold=1.8, print\_each=False)} still work?
    \item Why does the reordered call \texttt{run\_classifier\_pipeline(print\_each=False, threshold=1.8)} still work?
    \item When debugging, why is \textbf{Step Into} especially useful for studying arguments?
\end{enumerate}

\textbf{Micro-Exercise:} After students finish the tutorial, ask them to change the override threshold from \texttt{1.8} to \texttt{1.5}. Then ask them to compare the summary with the default run.
