\section{Positional Arguments}
Positional arguments are the most common way to pass values to functions. You've already been using them without explicitly realizing it since session 3 of our tutorial session.
 
In positional arguments, the order of the arguments provided in the function call directly corresponds to the order defined in the function signature.

\subsection{Loading Iris Data with a Single Positional Argument}

Take for example when we want to print a status message using the function \texttt{make\_print\_status}, which is defined to 
accept a single positional argument \texttt{status\_text}. When we call this function, we provide a string argument that describes the current status of our program.We are utilizing what is known as a positional argument, where the value we pass is matched to the parameter based on its position in the function definition. In this case, the string we provide will be assigned to the \texttt{status\_text} parameter when the function executes.



\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
make_print_status("Build dataset")

\end{minted}

The corresponding function definition is:

\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
def make_print_status(status_text):
    """Print a small status message.

    Args:
        status_text (str): A short message to show what the program is doing.
    """
    print(f"[STATUS] {status_text}")
\end{minted}

Here, the string "Build dataset" is passed as a positional argument to the \texttt{make\_print\_status} function. The function then uses this argument to print a status message indicating that the dataset is being built.



\subsection{Initializing initialize predictions with Two Positional Arguments}

Another example is the function \texttt{calculate\_accuracy}, defined with two positional parameters: \texttt{correct} and \texttt{total}.
When invoking this function, you must ensure that the arguments are provided in the correct sequence. Swapping these arguments inadvertently will cause unexpected behavior or errors in your program.

Thus, clear parameter naming and detailed documentation are essential to prevent such issues.
\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
accuracy = calculate_accuracy(correct, total)

\end{minted}



\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
def calculate_accuracy(correct, total):
    """Calculate accuracy percentage.

    Args:
        correct (int): Number of correct predictions.
        total (int): Number of processed samples.

    Returns:
        float: Accuracy percentage.
    """
    if total > 0:
        accuracy = (correct / total) * 100
    else:
        accuracy = 0.0

    return accuracy
    
\end{minted}






\subsection{ Update metrics Using Multiple Positional Arguments}
% update_metrics(metrics, y_pred, y_true)

Let's examine a function that requires multiple positional arguments, \texttt{update\_metrics}, which is responsible for rendering robots on the canvas:


\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
print_summary(correct, wrong, total, y_pred_list, accuracy)
\end{minted}


\begin{minted}[fontsize=\small, linenos, breaklines=true]{python}
def print_summary(correct, wrong, total, y_pred_list, accuracy):
    """Print the final results after the loop is finished.

    Args:
        correct (int): Number of correct predictions.
        wrong (int): Number of wrong predictions.
        total (int): Number of processed samples.
        y_pred_list (list): List of predicted labels.
        accuracy (float): Accuracy percentage.
    """
    print("\n=== session 4 Summary ===")
    print("Correct:", correct)
    print("Wrong:", wrong)
    print("Total:", total)
    print("Accuracy (%):", round(accuracy, 2))
    print("All predictions:", y_pred_list)     
\end{minted}

This function is built to accept FIVE  positional arguments (e.g., \texttt{correct}, \texttt{wrong}, \texttt{total}, \texttt{y\_pred\_list}, \texttt{accuracy}). 
When calling this function, you must provide all five arguments in the exact order they are defined in the function signature. For instance, if you accidentally swap the \texttt{correct} and \texttt{wrong} arguments, the output will be misleading and incorrect, as the function will interpret the values based on their position rather than their meaning. This highlights the importance of being cautious when using multiple positional arguments, as it can lead to confusion and errors if not handled carefully. Clear documentation and consistent naming conventions can help mitigate these risks, but it's crucial to always double-check the order of arguments when calling functions that rely on positional arguments.

\begin{tcolorbox}[title=Quiz]
Can you identify which other functions in our session 4 codebase also use multiple positional arguments?
\end{tcolorbox}


\subsection{Best Practices: Enhancing Code Clarity with Descriptive Names and Docstrings}

In scenarios with multiple positional arguments (2 or more), the significance of clear \texttt{docstring} and clear naming variables becomes even more importance. These practices help prevent confusion and errors, ensuring that anyone using the function (e.g., update\_metrics), understands what each argument represents and in what order they should be provided.

Yet, it's important to recognize that mistakes can still occur. Users might inadvertently supply incorrect input or mix up the order of arguments. So, I hope this highlight the important ensuring supplying the right value and order for correct operation of the function's operation.

\subsection{Self Reflection: Your Experience with Positional Arguments}
Nevertheless, either you realise it or not, but you actually have been working with the concept of positional arguments since the third session of our tutorials. So Kudos to you for that. Therefore, I did not design any specific task for this section, as I believe you already have a good grasp of how positional arguments work in \texttt{Python} functions.