"""session 3 student template: Iris rule classifier with lists, loops, and conditionals."""


# Session 2 continuity variables. No need to change these.
threshold = 2.0
feature_name = "petal_length"
positive_label = "setosa"
negative_label = "not_setosa"
label_key = "species"
correct = 0
wrong = 0
total = 0
y_pred_list = []


# Task 0: Replacing sequential variable with dict
"""
dict(id="flower2", 
sepal_length=5.1,
 ???=3.5, 
 # I deliberately drop two variables, and you should read the instructions in the template to add them back in. 
 petal_width=???, 
 species="setosa")
"""
# Dataset for session 3 sequential classifier (no functions yet)
dataset = [
    {
        "id": "flower1",
        "sepal_length": 5.1,
        "sepal_width": 3.5,
        "petal_length": 1.4,
        "petal_width": 0.2,
        "species": "setosa",
    },
    {
        "id": "flower2",
        "sepal_length": 4.9,
        "sepal_width": 3.0,
        "petal_length": 1.4,
        "petal_width": 0.2,
        "species": "setosa",
    },
    # add more samples...
]

# Task: Generate a for loop over the dataset.
"""
You should replace the ??? with the correct variable names to loop over the dataset and print out the id, petal_length, and species for each sample.

 the for loop, something like
for ??? in ???:
    print(sample["ID"], sample["petal_length"], sample["species"])
"""
# Decomment me
# for ??? in ???:
# print(sample["ID"], sample["petal_length"], sample["species"])

#   Task: Add if-else condition to classify each sample as setosa or not_setosa based on petal_length and threshold.

#   Task: Add code to convert true species to binary y_true label (setosa or not_setosa) and update the correct/wrong counters and prediction list.


# Hint: Everthing should be under the for loop, and you can use the same if-else condition to determine both y_pred and y_true.


# TODO 1: initialize counters and prediction list

print("\n=== Start session 3 Prediction Loop ===")


for sample in dataset:
    if sample[feature_name] < threshold:
        y_pred = positive_label
    else:
        y_pred = negative_label
    y_pred_list.append(y_pred)

    print(
        f"id={sample['id']} | pred={y_pred} | "
        f"petal_length={sample['petal_length']}"
    )


